---
title: "NBA Points Per Possession (PPP)"
author: "Andrew Tan"
output: html_document
---

Load data into dataframe `possessions`:
```{r}
setwd("/Users/andrewtan/Projects/Sports\ Analytics/basketball-analytics")
possessions <- read.csv("box_optical_poss.csv",
                 header = TRUE,
                 sep = ",")
possessions[is.na(possessions)] <- 0
possessions = possessions[as.character(possessions$possession.length) > "0:00", ]
```

## NBA Points Per Possession
Get the half court and transition PPP:
1. LA Lakers PPP
2. Opponent PPP
3. NBA Teams PPP

To retrieve a the PPP of any team, run `getTeamPPP(team_id)`
For example, Lakers' team PPP is `getTeamPPP(13)`

```{r TeamPPP}
getPPP = function(df=possessions) {
  n = nrow(df)
  points = sum(df$points)
  ppp = points / n
  return(ppp)
}
getTeamPPP = function(team_id, df=possessions) {
  team = df[df$team.global.id == team_id, ]
  ppp = getPPP(df=team)
  return(ppp)
}
```

```{r TestTeamPPP}
getTeamPPP(13)
```


## Transition PPP

Transition: Length of the possession is <= 8 seconds
To retrieve transition PPP of any team, run `getTransitionPPP(team_id)`
For example, Lakers' transition PPP is `getTransitionPPP(13)`

```{r transitionPPP}
getTransitionPPP = function(df=possessions, time="0:08") {
  transition = df[as.character(df$possession.length) <= time, ]
  ppp = getPPP(transition)
  return(ppp)
}

getTeamTransitionPPP = function(team_id, df=possessions, time="0:08") {
  team = df[df$team.global.id == team_id, ]
  ppp = getTransitionPPP(df=team)
  return(ppp)
}
```


## Half Court PPP

Half Court: Length of the possession is > 8 seconds
To retrieve half court PPP of any team, run `getHalfCourtPPP(team_id)`
For example, Lakers' half court PPP is `getHalfCourtPPP(13)`

```{r halfcourtPPP}
getHalfCourtPPP = function(df=possessions, time="0:08") {
  halfCourt = df[as.character(df$possession.length) > time, ]
  ppp = getPPP(halfCourt)
  return(ppp)
}

getTeamHalfCourtPPP = function(team_id, df=possessions, time="0:08") {
  team = df[df$team.global.id == team_id, ]
  ppp = getHalfCourtPPP(df=team)
  return(ppp)
}
```


## Assignment

Get the transition and halfcourt PPP of:
1. Lakers
2. Opponents
3. NBA Average

```{r assignment}
getOpponentsDF = function(team_id, df=possessions) {
  team = possessions[possessions$team.global.id == team_id, ]
  gameCodes = unique(team$gamecode)
  gamesDF = possessions[any(possessions$gamecode %in% gameCodes), ]
  opponentsDF = gamesDF[gamesDF$team.global.id != team_id, ]
  return(opponentsDF)
}

# Get Lakers Opponents
lakersOpponentsDF = getOpponentsDF(13)


# Print Values
paste("Lakers Transition PPP: ", getTeamTransitionPPP(13))
paste("Lakers Half Court PPP: ", getTeamHalfCourtPPP(13))

paste("Opponent Transition PPP: ", getTransitionPPP(lakersOpponentsDF))
paste("Opponent Half Court PPP: ", getHalfCourtPPP(lakersOpponentsDF))

paste("NBA Transition PPP: ", getTransitionPPP(possessions))
paste("NBA Half Court PPP: ", getHalfCourtPPP(possessions))
```
